Supplemental Information A: Full Annotated Analysis and Reproducible Code

### 2022-11-27 methods-and-results.Rmd
:construction: pca code

Rdimtools::do.udfs(scaled_data, k = 5)

res.pca <- prcomp(scaled_data, scale = FALSE)
fviz_eig(res.pca, addlabels = TRUE)
summary(res.pca)

var <- get_pca_var(res.pca)
corrplot::corrplot(var$contrib[,1:10], is.corr=FALSE) 
fviz_contrib(res.pca, choice = "var", axes = 1, top = 50)
fviz_contrib(res.pca, choice = "var", axes = 2, top = 50)

res.desc <- FactoMineR::dimdesc(res.pca, axes = c(1,2), proba = 0.05)
# Description of dimension 1
res.desc$Dim.1

@@

pca.out <- as.data.frame(res.pca$x[,1:3])

For the

\subsubsection{Feature clustering}

>>>>

res.pca <- prcomp(scaled_data, scale. = FALSE)
summary(res.pca)

res.varExplained <- data.frame(
  Explained_Variance = cumsum(summary(res.pca)[["importance"]]['Proportion of Variance',]),
  Components = seq(1, length(cumsum(summary(res.pca)[["importance"]]['Proportion of Variance',])), 1)
)
fviz_eig(res.pca, addlabels = TRUE)
ggplot(data = res.varExplained, aes(x = Components, y = Explained_Variance)) +
  geom_point() +
  geom_line() +
  theme_Publication()

pca_cutoff <- min(which(cumsum(summary(res.pca)[["importance"]]['Proportion of Variance',]) >= 0.8))

@@

pca.out <- as.data.frame(res.pca$x[,1:pca_cutoff])
```

For the feature reduction, we chose the very general \textit{principal component analysis} (PCA). While some of the more tailor-made feature selection algorithms can be more accurate and meaningful in reducing the feature dimensionality, PCAs are well-established and can broadly be applied in an automatize manner. 80% rule of thumb

we select the first `r pca_cutoff` principal components because they explain 80\% of the variance in the original `r ncol(scaled_data)` features (reducing the dimensions by `r format(round((1-pca_cutoff/ncol(scaled_data))*100, 2), nsmall=2)`\%).

\subsubsection{Feature clustering}
```{r featureClustering, include=FALSE}
# load libraries
library(factoextra)
library(NbClust)

# prep data
scaled_data <- pca.out %>% scale

# Elbow method
fviz_nbclust(scaled_data, kmeans, method = "wss") +
  geom_vline(xintercept = 2, linetype = 2)+
  labs(subtitle = "Elbow method")
# Silhouette method
fviz_nbclust(scaled_data, kmeans, method = "silhouette")+
  labs(subtitle = "Silhouette method")
# Gap statistic [takes for ever]
# nboot = 50 to keep the function speedy.
# recommended value: nboot= 500 for your analysis.
# Use verbose = FALSE to hide computing progression.
# set.seed(123)
# fviz_nbclust(scaled_data, kmeans, nstart = 100,  method = "gap_stat", nboot = 500)+
#   labs(subtitle = "Gap statistic method")
```

### 2022-11-27 methods-and-results.tex
:construction: pca code

>>>>

For the feature reduction, we chose the very general
\textit{principal component analysis} (PCA). While some of the more
tailor-made feature selection algorithms can be more accurate and
meaningful in reducing the feature dimensionality, PCAs are
well-established and can broadly be applied in an automatize manner.
80\% rule of thumb

we select the first 27 principal components because they explain 80\% of
the variance in the original 72 features (reducing the dimensions by
62.50\%).
